---
title: "Case Study 2"
author: "Justin Cleveland"
date: "4/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction  
Code below was developed to analyze employee data for the purpose of predicting employee turnover. The R code goes through a series of step to import raw data, transform, visualize, and develop prediction based models. The primary goal is to identify factors that lead to employee attrition, however there is additional code developed to predict an employee's monthly income. 
  
You should identify the top three factors that contribute to turnover (backed up by evidence provided by analysis). There may or may not be a need to create derived attributes/variables/features.
  
Inaddition The business is also interested in learning about any job role specific trends that may exist in the data set (e.g., “Data Scientists have the highest job satisfaction”). You can also provide any other interesting trends and observations from your analysis. 
  
Classification Model to predict attriton (Plus top three factors, and at least 60% sensitivity and specificity (60 each = 120 total))
Prediction Model of Monthly incomes (Lowest RMSE model)
Jobe role trends
  
  
### Load required R Libraries
```{r,message=FALSE,warning=FALSE}
library(tidyverse) #Loads core packages including:
#ggplot2, for data visualisation
#dplyr, for data manipulation
#tidyr, for data tidying
#readr, for data import
#purrr, for functional programming
#tibble, for tibbles, a modern re-imagining of data frames
#stringr, for strings
#forcats, for factors
library(GGally) #Used to plot relationship of variables in MLR model
library(ggfortify) #Used to plot MLR residuals and Q-Q plot
library(Metrics) #Used to calculate mse 
library(gridExtra) #Used to a more complex graphics to ggplot, grid.arrange function
library(Hmisc) #Used to build correlation matrix table
library(plyr) #Used for data analysis and tidy data
library(caret) #Used to split data into training and testing
library(e1071) #Used for naiveBayes model
library(RCurl) #Used to read data using GetURL fuction
library(corrplot) #Used to build correlation matrix
```

### Import Data 
Code below will read data into R from Github
```{r}
# Read in data for analysis
Data = read.csv(text=getURL("https://raw.githubusercontent.com/JustinCleveland/CaseStudy2DDS/master/CaseStudy2-data.csv"), header = TRUE,   na.strings =c('', 'NA'))
Dataset = Data
Validation_Attrition = read.csv("https://raw.githubusercontent.com/JustinCleveland/CaseStudy2DDS/master/CaseStudy2CompSet No Attrition.csv",header = TRUE,na.strings=c("","NA"))
Validation_Salary = read.csv("https://raw.githubusercontent.com/JustinCleveland/CaseStudy2DDS/master/CaseStudy2CompSet No Salary.csv",header = TRUE,na.strings=c("","NA"))

```

### Transform Data
#### Change factor variables to ordinal variables

```{r}
#Attrition
Dataset$Att <-            ifelse(grepl("*No*",Dataset$Attrition) , "0", "1")
                          Dataset$Att = as.integer(Dataset$Att)

#Gender
Dataset$Gender <-         ifelse(grepl("*Female*",Dataset$Gender) , "0", "1")
                          Dataset$Gender = as.integer(Dataset$Gender)

#Overtime
Dataset$OverTime <-       ifelse(grepl("*No*",Dataset$OverTime) , "0", "1")
                          Dataset$OverTime = as.integer(Dataset$OverTime)

#BusinessTravel                                              
Dataset$BusinessTravel <- ifelse(grepl("*Non-Travel*", Dataset$BusinessTravel), "0", 
                          ifelse(grepl("*Travel_Frequently*", Dataset$BusinessTravel) , "1", "2"))
                          Dataset$BusinessTravel = as.integer(Dataset$BusinessTravel)

#Department
Dataset$Department <-     ifelse(grepl("*Human Resources*", Dataset$Department ), "0", 
                          ifelse(grepl("*Research & Development*", Dataset$Department) , "1", "2"))
                          Dataset$Department = as.integer(Dataset$Department )

#MaritalStatus
Dataset$MaritalStatus <-  ifelse(grepl("*Divorced*", Dataset$MaritalStatus), "0", 
                          ifelse(grepl("*Married*", Dataset$MaritalStatus) , "1", "2"))
                          Dataset$MaritalStatus = as.integer(Dataset$MaritalStatus)

#EducationField
Dataset$EducationField <- ifelse(grepl("*Human Resources*", Dataset$EducationField), "0", 
                          ifelse(grepl("*Life Sciences*", Dataset$EducationField) , "1", 
                          ifelse(grepl("*Marketing*", Dataset$EducationField) , "2", 
                          ifelse(grepl("*Medical*", Dataset$EducationField) , "3", 
                          ifelse(grepl("*Other*", Dataset$EducationField) , "4", 
                                 "5")))))
                          Dataset$EducationField = as.integer(Dataset$Department)

#JobRole
Dataset$JobRole <-        ifelse(grepl("*Healthcare Representative*", Dataset$JobRole), "0", 
                          ifelse(grepl("*Human Resources*", Dataset$JobRole) , "1", 
                          ifelse(grepl("*Laboratory Technician*", Dataset$JobRole) , "2", 
                          ifelse(grepl("*Manager*", Dataset$JobRole) , "3", 
                          ifelse(grepl("*Manufacturing Director*", Dataset$JobRole) , "4", 
                          ifelse(grepl("*Research Director*", Dataset$JobRole) , "5", 
                          ifelse(grepl("*Research Scientist*", Dataset$JobRole) , "6", 
                          ifelse(grepl("*Sales Executive*", Dataset$JobRole) , "7", 
                          "8"))))))))
                          Dataset$JobRole = as.integer(Dataset$JobRole)
```

#### Create Table with variables that have more than one level  
  
Excludes variables: EmployeeCount,EmployeeNumber, Over18, StandardHours

```{r}

Dataset = Dataset %>% select(ID,Attrition,Att,Age,BusinessTravel,DailyRate,Department,DistanceFromHome,Education,EducationField,EnvironmentSatisfaction,Gender,HourlyRate,JobInvolvement,JobLevel,JobRole,JobSatisfaction,MaritalStatus,MonthlyIncome,MonthlyRate,NumCompaniesWorked,OverTime,PercentSalaryHike,PerformanceRating,RelationshipSatisfaction,StockOptionLevel,TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager)

```

#### Create Test Dataset and Training Dataset using transformed variables 

```{r,message=FALSE}
set.seed(4)
Dataset_Index = sample(seq(1:length(Dataset$ID)),round(.7*length(Dataset$ID)))
Dataset_Train = Dataset[Dataset_Index,]
Dataset_Test = Dataset[-Dataset_Index,]
```

### Visualize  


```{r,message=FALSE}
#Bar Plot of Employee Attrition
Data %>% 
  ggplot(aes(x=Attrition)) +  
  geom_bar() + 
  ggtitle("Count of Employee Attrition") + xlab("Attrition") + ylab("Count") 

#Group TotalWorking Years and Monthly Income into buckets
Data <- Data %>% mutate(TotWorkingYears=cut(TotalWorkingYears, breaks=c(-Inf, 6, 11, 16,21,31,Inf), labels=c(">5","6-10","11-15","16-20","21-30","<30")))
Data <- Data %>% mutate(MoIncome=cut(MonthlyIncome, breaks=c(-Inf, 5000, 10000, 15000,Inf), labels=c(">$5k","$5K-$10K","$11K-$15k","<$16k")))


Data %>% select(Attrition,MonthlyIncome) %>%
  ggplot(aes(x=Attrition,y=MonthlyIncome,fill=Attrition)) +  
  geom_boxplot(outlier.shape=NA) +
  geom_point(aes(fill=Attrition) ,position=position_jitterdodge(jitter.width = .1)) +
  ggtitle("Attrition by MonthlyIncome") + xlab("Attrition") + ylab("MonthlyIncome")


OT = Data %>% select(Attrition,MonthlyIncome,OverTime) %>%
  ggplot(aes(x=OverTime,y=MonthlyIncome,fill=Attrition)) +  
  geom_boxplot(outlier.shape=NA) +
  geom_point(aes(fill=Attrition) ,position=position_jitterdodge(jitter.width = .1)) +
  ggtitle("Attrition by MonthlyIncome and OverTime") + xlab("OverTime") + ylab("MonthlyIncome") 

TW = Data %>% select(Attrition,MonthlyIncome,TotWorkingYears) %>%
  ggplot(aes(x=TotWorkingYears,y=MonthlyIncome,fill=Attrition)) +  
  geom_boxplot(outlier.shape=NA) +
  geom_point(aes(fill=Attrition) ,position=position_jitterdodge(jitter.width = .1)) +
  ggtitle("Attrition by MonthlyIncome and TotalWorkingYears") + xlab("Total Working Years") + ylab("MonthlyIncome") 

grid.arrange(OT, TW, ncol=2)

```


#### Identify variables that are correlatted
Custom function to format the correlation matrix to help identify variables correlatted with attrition  
Source for custom function code: https://rstudio-pubs-static.s3.amazonaws.com/240657_5157ff98e8204c358b2118fa69162e18.html

```{r,message=FALSE}
#Custom function to create table of correlated variables
flat_cor_mat <- function(cor_r, cor_p){
  #This function provides a simple formatting of a correlation matrix
  #into a table with 4 columns containing :
    # Column 1 : row names (variable 1 for the correlation test)
    # Column 2 : column names (variable 2 for the correlation test)
    # Column 3 : the correlation coefficients
    # Column 4 : the p-values of the correlations
  cor_r <- rownames_to_column(as.data.frame(cor_r), var = "row")
  cor_r <- gather(cor_r, column, cor, -1)
  cor_p <- rownames_to_column(as.data.frame(cor_p), var = "row")
  cor_p <- gather(cor_p, column, p, -1)
  cor_p_matrix <- left_join(cor_r, cor_p, by = c("row", "column"))
  cor_p_matrix
}

#Create plot of variables
corrplot(cor(Dataset[,c(3:33)]), method = "ellipse")

#Create table showing the correlation coeffient with attrition
CorrMatrix <- rcorr(as.matrix(Dataset[,c(3:33)]))
CorrTable = flat_cor_mat(round(CorrMatrix$r,3), round(CorrMatrix$P,3))
CorrTable  %>% 
  arrange(p) %>%
  filter(column=='Att') %>%
  select(row,cor,p)

```

### Model 

#### Create Naïve Bayes Classification Model to predict employee attrition

```{r,message=FALSE}
# Use original data including factors and numerical data
NB_Dataset = Data %>% select(ID,Attrition,Age,BusinessTravel,DailyRate,Department,DistanceFromHome,Education,EducationField,EnvironmentSatisfaction,Gender,HourlyRate,JobInvolvement,JobLevel,JobRole,JobSatisfaction,MaritalStatus,MonthlyIncome,MonthlyRate,NumCompaniesWorked,OverTime,PercentSalaryHike,PerformanceRating,RelationshipSatisfaction,StockOptionLevel,TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager)
#str(NB_Dataset)

# Create NB Model Test and Training Datasets
NB_Dataset_Index = sample(seq(1:length(NB_Dataset$ID)),round(.7*length(NB_Dataset$ID)))
NB_Dataset_Train = NB_Dataset[NB_Dataset_Index,]
NB_Dataset_Test = NB_Dataset[-NB_Dataset_Index,]

NB_Attrition_Model = naiveBayes(NB_Dataset_Train[,c(3:32)],as.factor(NB_Dataset_Train$Attrition),laplace = 1)
confusionMatrix(table(predict(NB_Attrition_Model,NB_Dataset_Test[,c(3:32)]),as.factor(NB_Dataset_Test$Attrition)))
```


#### Second pass NB model
Build model with highly correlated variables with attrition

```{r,message=FALSE, warning=FALSE}
#Final variables selected for NB model which are based on the lowest pvalue in the correlation table 
NB_Dataset_2 = Dataset %>% select(ID,Att,Attrition,JobInvolvement, JobLevel, MaritalStatus, OverTime, TotalWorkingYears, MonthlyIncome, YearsInCurrentRole)
str(NB_Dataset_2)

#Create table of variables correlated with attrition
CM_Dataset = NB_Dataset_2
CorrMatrix <- rcorr(as.matrix(CM_Dataset[,c(2,4:10)]))
CorrTable <- flat_cor_mat(round(CorrMatrix$r,4), round(CorrMatrix$P,06))
CorrTable[CorrTable$column=='Att',]

#Create plot of variables correlated with attrition
corrplot(cor(CM_Dataset[,c(2,4:10)]), method = "ellipse")

set.seed(6)
NB_Dataset_Index_2 = sample(seq(1:length(NB_Dataset_2$ID)),round(.7*length(NB_Dataset_2$ID)))
NB_Dataset_Train_2 = NB_Dataset_2[NB_Dataset_Index_2,]
NB_Dataset_Test_2 = NB_Dataset_2[-NB_Dataset_Index_2,]

#Final NB Model
NB_Attrition_Model_2 = naiveBayes(NB_Dataset_Train_2[,c(3:10)],as.factor(NB_Dataset_Train_2$Attrition),laplace = 1)
confusionMatrix(table(predict(NB_Attrition_Model_2,NB_Dataset_Test_2[,c(3:10)]),as.factor(NB_Dataset_Test_2$Attrition)))

#Top three variables that lead to attrition JobInvolvement, MaritalStatus, OverTime
model = train(Attrition~.,data = NB_Dataset,method = 'nb', trControl=trainControl(method='cv',number=10))
x = varImp(model)
Imp = rownames_to_column(as.data.frame(varImp(model)$importance))
Imp %>% ggplot(aes(x = reorder(rowname, Yes), y = Yes)) +
  geom_bar(stat = 'identity') + 
  coord_flip() + 
  ggtitle('Factors Influencing Employee Attrition') + 
  ylab('% of Importance') +
  xlab('')

#Change factors to ordinal variables
NB_Validation = Validation_Attrition
#Overtime
NB_Validation$OverTime <-       ifelse(grepl("*No*",NB_Validation$OverTime) , "0", "1")
                          NB_Validation$OverTime = as.integer(NB_Validation$OverTime)

#MaritalStatus
NB_Validation$MaritalStatus <-  ifelse(grepl("*Divorced*", NB_Validation$MaritalStatus), "0", 
                          ifelse(grepl("*Married*", NB_Validation$MaritalStatus) , "1", "2"))
                          NB_Validation$MaritalStatus = as.integer(NB_Validation$MaritalStatus)

NB_Validation$Attrition = predict(NB_Attrition_Model_2, newdata = NB_Validation)
Case2PredictionsCLEVELAND_Attrition = NB_Validation %>% select(ID,Attrition)

write.csv(Case2PredictionsCLEVELAND_Attrition, file = "C:/Users/cleve/OneDrive/Documents/R/R Files/SMU/Doing Data Science/Case_Study_2/Case2PredictionsCLEVELAND_Attrition.csv")

```


## Predict monthly income using regression model  
You must provide a model that will attain a RMSE < $3000 for the training and the validation set.  
  

#### Transform and Visualize MLR Data  
Create MLR Dataset with all variables from the training and test dataset
```{r,message=FALSE}

MLR_Train = Dataset_Train %>% select(ID,Att,Age,BusinessTravel,DailyRate,Department,DistanceFromHome,Education,EducationField,EnvironmentSatisfaction,Gender,HourlyRate,JobInvolvement,JobLevel,JobRole,JobSatisfaction,MaritalStatus,MonthlyIncome,MonthlyRate,NumCompaniesWorked,OverTime,PercentSalaryHike,PerformanceRating,RelationshipSatisfaction,StockOptionLevel,TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager)

MLR_Test = Dataset_Test %>% select(ID,Att,Age,BusinessTravel,DailyRate,Department,DistanceFromHome,Education,EducationField,EnvironmentSatisfaction,Gender,HourlyRate,JobInvolvement,JobLevel,JobRole,JobSatisfaction,MaritalStatus,MonthlyIncome,MonthlyRate,NumCompaniesWorked,OverTime,PercentSalaryHike,PerformanceRating,RelationshipSatisfaction,StockOptionLevel,TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager)
```

##### First Pass MLR Model using all variables

```{r,message=FALSE}
# Fit Model with all variables from MLR_Train Dataset
MLR_FitALL <- lm(formula = MonthlyIncome ~ . -ID,data=MLR_Train)
summary(MLR_FitALL)

#Calculate RMSE using MLR_FitALL model on Test Data
MLR_Test$PredictMI_ALL = predict(MLR_FitALL, newdata = MLR_Test)
rmse(MLR_Test$MonthlyIncome,MLR_Test$PredictMI_ALL)

```

#### Second and Third Pass MLR Models 

```{r,message=FALSE, warning=FALSE}
#MLR Correlation Plot
corrplot(cor(MLR_Train), method = "ellipse")

#MLR Correlation Table
MLR_CorrMatrix <- rcorr(as.matrix(MLR_Train))
MLR_CorrTable <- flat_cor_mat(round(MLR_CorrMatrix$r,4), round(MLR_CorrMatrix$P,5))
MLR_CorrTable[MLR_CorrTable$column=='MonthlyIncome',]

#Second Pass Model with variables correlation to monthly income
MLR_FitTEST <- lm(formula = MonthlyIncome ~ Age + JobLevel + MaritalStatus + TotalWorkingYears,data=MLR_Test)
summary(MLR_FitTEST)

#Third Pass Model using backward elimination Variable Selection process
t=step(MLR_FitALL,direction = "backward")

MLR_FitBACKWARDS <- lm(formula = MonthlyIncome ~ YearsWithCurrManager + BusinessTravel + YearsAtCompany + DistanceFromHome + EnvironmentSatisfaction + JobRole +  TotalWorkingYears + Department + JobLevel, data = MLR_Test)
summary(MLR_FitBACKWARDS)
```

#### Calculate RMSE with all models on the test data, select final model, visualize assumptions


```{r,message=FALSE, warning=FALSE}
#Calculate RMSE using MLR_FitALL model on Test Data
MLR_Test$PredictMI_ALL = predict(MLR_FitALL, newdata = MLR_Test)
rmse(MLR_Test$MonthlyIncome,MLR_Test$PredictMI_ALL)

#Calculate RMSE using MLR_FitTEST model on Test Data
MLR_Test$MLR_FitTEST = predict(MLR_FitTEST, newdata = MLR_Test)
rmse(MLR_Test$MonthlyIncome,MLR_Test$MLR_FitTEST)

#Calculate RMSE using MLR_FitBACKWARDS model on Test Data
MLR_Test$MLR_FitBACKWARDS = predict(MLR_FitBACKWARDS, newdata = MLR_Test)
rmse(MLR_Test$MonthlyIncome,MLR_Test$MLR_FitBACKWARDS)

#Residual and Q-Q plot
autoplot(MLR_FitBACKWARDS,1:2)

#Plot MLR variables
MLR_Train %>% select(MonthlyIncome,YearsWithCurrManager,BusinessTravel,YearsAtCompany,DistanceFromHome,EnvironmentSatisfaction,JobRole, TotalWorkingYears,Department,JobLevel) %>% ggpairs()

```

#### Run final model on validation dataset and write csv file

```{r,message=FALSE}
#### Change factor variables to ordinal variables for validation dataset
MLR_Validation = Validation_Salary
MLR_Validation = Validation_Salary %>% select(ID,YearsWithCurrManager,BusinessTravel,YearsAtCompany,DistanceFromHome,EnvironmentSatisfaction,JobRole, TotalWorkingYears,Department,JobLevel)

#BusinessTravel                                              
MLR_Validation$BusinessTravel <- ifelse(grepl("*Non-Travel*", MLR_Validation$BusinessTravel), "0", 
                          ifelse(grepl("*Travel_Frequently*", MLR_Validation$BusinessTravel) , "1", "2"))
                          MLR_Validation$BusinessTravel = as.integer(MLR_Validation$BusinessTravel)

#Department
MLR_Validation$Department <-     ifelse(grepl("*Human Resources*", MLR_Validation$Department ), "0", 
                          ifelse(grepl("*Research & Development*", MLR_Validation$Department) , "1", "2"))
                          MLR_Validation$Department = as.integer(MLR_Validation$Department )

#JobRole
MLR_Validation$JobRole <-        ifelse(grepl("*Healthcare Representative*", MLR_Validation$JobRole), "0", 
                          ifelse(grepl("*Human Resources*", MLR_Validation$JobRole) , "1", 
                          ifelse(grepl("*Laboratory Technician*", MLR_Validation$JobRole) , "2", 
                          ifelse(grepl("*Manager*", MLR_Validation$JobRole) , "3", 
                          ifelse(grepl("*Manufacturing Director*", MLR_Validation$JobRole) , "4", 
                          ifelse(grepl("*Research Director*", MLR_Validation$JobRole) , "5", 
                          ifelse(grepl("*Research Scientist*", MLR_Validation$JobRole) , "6", 
                          ifelse(grepl("*Sales Executive*", MLR_Validation$JobRole) , "7", 
                          "8"))))))))
                          MLR_Validation$JobRole = as.integer(MLR_Validation$JobRole)

MLR_Validation$MonthlyIncome = predict(MLR_FitBACKWARDS, newdata = MLR_Validation)
Case2PredictionsCLEVELAND_Salary = MLR_Validation %>% select(ID,MonthlyIncome)

write.csv(Case2PredictionsCLEVELAND_Salary, file = "C:/Users/cleve/OneDrive/Documents/R/R Files/SMU/Doing Data Science/Case_Study_2/Case2PredictionsCLEVELAND_Salary.csv")


```
